---
title: "Cleaning data Morphometrics"
author: "Hannah"
date: "January 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Loading required packages

```{r message=FALSE, warning=FALSE}
library(plyr)
library(tidyverse)
library(cowplot)
library(rptR)
library(dplyr)
```

## Importing raw data
This a datafile that contains all measurements taken even the ones that were off. There is duplication of samples that can be used for repeatability calcualtions.

Step 1) Import
Step 2) Convert from wide to long format
Step 3) Remove NA, retain only AF and AM values, and remove all measures smaller than 10

```{r data}
setwd("D:/Rfiles/Tick_Count_gen")
data<-read_csv("data/morph_raw_2.csv")
data1<-gather(data,4:37, key = "measure",value="value")
data1<-filter(data1,Age=="AF" | Age=="AM")
data1<-filter(data1,value>10)
```

##Graphing histogram of all measures broken down by age

```{r graph histoALL}
plot0<- ggplot(aes((value)), data = data1)+
  geom_histogram()+
  scale_x_log10()+
  geom_vline(xintercept=10,color="red")+
  facet_grid(Age~.)
plot0
```
##A function for added min and max box plots to plot

```{r function min max}
min.mean.sd.max <- function(x) {
  r <- c(min(x), mean(x) - sd(x), mean(x), mean(x) + sd(x), max(x))
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r
}
```

##Graphing all measures summarized by age.

```{r graphing all}
plot1 <- ggplot(aes(y = (value), x=measure), data = data1)
plot1 <- plot1 + stat_summary(fun.data = min.mean.sd.max, geom = "boxplot") + 
  geom_jitter()+
  scale_y_log10()+
  facet_grid(Age~.,scales="free")
plot1
```


##Cleaning up the data to remove crazy values.

Step 1) Calculation of mean based on current values (mean of measure/by age) and append to table
Step 2) Add the percent difference of value from mean. This is just a measure of how differnet the values are
Step 3) Values that differ by an order of 10 from the mean are assumed to be missing a decimal place and are multipled/divided by 10 accordingly.
Step 4) Repeat step 1-4 assuming some values will be missed because mean will be influenced by scewed numbers

```{r subsetting}
#Step one
averages<- ddply(data1,c("Age","measure"),summarise,
          mean = mean(value,na.rm=TRUE))
data2<-left_join(data1,averages,by = c("Age","measure"))

#Step two
data2<-data2 %>% mutate(PEC = value/mean)

##an aside
data2d<-data2

#Step three
data2$value <-if_else(data2$PEC<0.1,data2$value*10,
                        if_else(data2$PEC>10,data2$value*0.1,data2$value))
data2<-select(data2,1:5)

#Steo four
averages<- ddply(data2,c("Age","measure"),summarise,
          mean = mean(value,na.rm=TRUE))
data2<-left_join(data1,averages,by = c("Age","measure"))
data2<-data2 %>% mutate(PEC = value/mean)

data2$value <-if_else(data2$PEC<0.1,data2$value*10,
                        if_else(data2$PEC>10,data2$value*0.1,data2$value))
dataPEC<-select(data2,1:4,7)
dataPEC2<-spread(dataPEC,measure,PEC)
data2<-select(data2,1:5)
```

My aside
```{r graphing all}
data2d$Date<-dmy(data2d$Date)
plot1a <- ggplot(aes(y = PEC, x=Date), data = data2d)
plot1a <- plot1 + geom_point(aes(colour=measure))+
  #scale_y_log10()+
  facet_grid(Age~.,scales="free")
plot1a
```
Quick graphing histo of percentage of mean values

```{r graph histoALL}
plotPEC<- ggplot(aes((PEC)), data = dataPEC)+
  geom_histogram()+
  geom_vline(xintercept=1,color="red")+
  facet_grid(Age~.)
plotPEC
```

##More cleaning

This is a more rigorous cleaning of the data.

Step 1)Calculate the average and sd
Step 2)Determine high and low cut off values. Currently set at High=3SD and Low=3SD
Step 3)Remove duplicate values taking the mean of two measurments
Steo

```{r graphing2}
#Step one
averages2<- ddply(data2,c("Age","measure"),summarise,
          mean = mean(value,na.rm=TRUE),
          sd = sd(value,na.rm = TRUE))

#Step two
averages2<- averages2 %>% mutate(High = mean+(3*sd), Low = mean-(3*sd))
data2<-left_join(data2,averages2,by = c("Age","measure"))
data2<-filter(data2,value>Low)
data2<-filter(data2,value<High)

#Step three
data2x<-select(data2,1:5)
data2x<-ddply(data2x,c("Tick_vial", "Age","measure"),summarise,
          value = mean(value,na.rm=TRUE))
```

#A quick check of data quality

Look at varianence and mean of each value
```{r graphing2}
variance<- ddply(data2x,c("Age","measure"),summarise,
          mean = mean(value,na.rm=TRUE),
          sd = sd(value,na.rm = TRUE),
          variance = sd(value,na.rm = TRUE)/mean(value,na.rm=TRUE))
```

#Transform cleaned data from long to wide

This involves breaking the data up into AM and AF as well

```{r data transform}
data3<-spread(data2x,measure,value)
data3F<-filter(data3, Age =="AF")
data3M<-filter(data3, Age =="AM")
```

```{r counting NA}
data3F$na_count <- apply(is.na(data3F), 1, sum)
data3M$na_count <- apply(is.na(data3M), 1, sum)
data3F<-filter(data3F,na_count<6)
data3M<-filter(data3M,na_count<6)
sapply(data3F, function(y) sum(length(which(is.na(y)))))
sapply(data3M, function(y) sum(length(which(is.na(y)))))

data3F<-select(data3F,-APL,-APW,-DPL,-DPW,-na_count)
data3M<-select(data3M,-CHL,-CHW,-na_count)
sapply(data3F, function(y) sum(length(which(is.na(y)))))
sapply(data3M, function(y) sum(length(which(is.na(y)))))

for(i in 1:ncol(data3F)){
  data3F[is.na(data3F[,i]), i] <- mean(data3F[,i], na.rm = TRUE)
}
for(i in 1:ncol(data3M)){
  data3M[is.na(data3M[,i]), i] <- mean(data3M[,i], na.rm = TRUE)
}
```

```{r print to csv}
write.csv(data3F, "data/data3F.csv")
write.csv(data3M, "data/data3M.csv")
```




