---
title: "Borrelia Serological Analysis"
output: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up the file

Setting the working directory, loading packages, and loading data.

```{r}
setwd("D:/Rfiles/TickPrev") 
data<-read.csv("data/ser.csv",header=TRUE)
library(MASS)
library(car)
```

## Data Check

Just to ensure that everything loaded properly. I look at the head() so the top five rows. I also look at the struture which should let me know if anything imported funny and which rows are catagorical or continous.

I also changed to variables to factors so that they run properly in the GLM

```{r}
head(data)
str(data)

data$pos<-factor(data$pos)
data$Year<-factor(data$Year)
```

## GLM time

Loading the models.

I am looking at the rate of positive based on serology related to species of collection, location/colony, and year.

This is binomial because it is postive/negative data

Finally print the AIC values to determine the 'top' model.

```{r}
m1 <- with(data, glm(pos ~ Species, family = binomial))
m2 <- with(data, glm(pos ~ Loc, family = binomial))
#m3 <- with(data, glm(pos ~ Year, family = binomial))
m4 <- with(data, glm(pos ~ Species+Loc, family = binomial))
#m5 <- with(data, glm(pos ~ Species+Year, family = binomial))
#m6 <- with(data, glm(pos ~ Loc+Year, family = binomial))
#m7 <- with(data, glm(pos ~ Loc+Year+Species, family = binomial))

AIC(m1,m2,m4)
```

It is important to bring the loglik values too.

```{r}
logLik(m1)
logLik(m2)
logLik(m4)

```

I always take a look at the top model to see what is significant and ensure nothing funky has happened

```{r}
summary(m1)
Anova(m2,3)
```

## Comparisons in top model

Now this is questionable as it is mixing multiple statitical approches but this is what was wanted in my thesis.

This is how to look at the between catagory comparisons using Tukey test in the top model

```{r}
library(multcomp)
summary(glht(m1, mcp(Species="Tukey")))
```

